When Subgraph Isomorphism is Really Hard, and Why This Matters for Graph Databases

### 2018-02-14 paper/paper.tex
Clarify heuristics (review 1)

This work builds upon the phase transition phenomena observed in satisfiability and graph colouring
problems first described by \citet{DBLP:conf/ijcai/CheesemanKT91} and
\citet{DBLP:conf/aaai/MitchellSL92}. For subgraph isomorphism in Erd\H{o}s-R\'enyi random graphs, we
identify three relevant control parameters: we can independently alter the edge probability of the
pattern graph, the edge probability of the target graph, and the relative orders (number of
vertices) of the pattern and target graphs.  For non-induced isomorphisms, with the correct choice
of parameters we see results very similar to those observed with boolean satisfiability problems:
there is a phase transition (whose location we can predict) from satisfiable to unsatisfiable, and
we see a solver-independent complexity peak near this phase transition. Additionally, understanding
this behaviour helps us to design better variable- and value-ordering heuristics---this is the
second contribution of this paper.

@@

The Glasgow and LAD solvers are constraint programming inspired algorithms (although they use
dedicated data structures, and do not use a constraint programming toolkit). Constraint programming
is a field concerned with solving \emph{constraint satisfaction} problems. In such a problem, we
have a set of \emph{variables}, each of which has a \emph{domain} of \emph{values}. We also have a
set of \emph{constraints}. The objective is to assign each variable a value from its domain, whilst
respecting every constraint. This is done using a combination of \emph{inference}, which eliminates
infeasible values from domains, and backtracking search. This search is driven by
\emph{variable-ordering} heuristics, which select which unassigned variable to branch on, and
\emph{value-ordering} heuristics, which select which value to try giving that variable. If at any
point during search a \emph{domain wipeout} occurs (that is, if every value is removed from any
domain), the search will backtrack immediately.

>>>>

This work builds upon the phase transition phenomena observed in satisfiability and graph colouring
problems first described by \citet{DBLP:conf/ijcai/CheesemanKT91} and
\citet{DBLP:conf/aaai/MitchellSL92}. For subgraph isomorphism in Erd\H{o}s-R\'enyi random graphs, we
identify three relevant control parameters: we can independently alter the edge probability of the
pattern graph, the edge probability of the target graph, and the relative orders (number of
vertices) of the pattern and target graphs.  For non-induced isomorphisms, with the correct choice
of parameters we see results very similar to those observed with boolean satisfiability problems:
there is a phase transition (whose location we can predict) from satisfiable to unsatisfiable, and
we see a solver-independent complexity peak near this phase transition. Additionally, understanding
this behaviour helps us to design better search strategies, by improving variable- and
value-ordering heuristics---this is the second contribution of this paper.

@@

The Glasgow and LAD solvers are constraint programming inspired algorithms (although they use
dedicated data structures, and do not use a constraint programming toolkit). Constraint programming
is a field concerned with solving \emph{constraint satisfaction} problems. In such a problem, we
have a set of \emph{variables}, each of which has a \emph{domain} of \emph{values}. We also have a
set of \emph{constraints}. The objective is to assign each variable a value from its domain, whilst
respecting every constraint. This is done using a combination of \emph{inference}, which eliminates
infeasible values from domains, and backtracking search. This search is driven by
\emph{variable-ordering} heuristics, which select which unassigned variable to branch on, and
\emph{value-ordering} heuristics, which select which value to try giving that variable. That is,
variable-ordering selects a pattern vertex, whilst value-ordering tells us the order in which to try
different candidate target vertices for that pattern vertex. If at any point during search a
\emph{domain wipeout} occurs (that is, if every value is removed from any domain), the search will
backtrack immediately.
